---
title: "Journal (reproducible report)"
author: "Nariman Seifinezhad"
date: "2020-11-05"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    toc_depth: 3
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

**IMPORTANT:** You can delete everything in here and start fresh. You might want to start by not deleting anything above this line until you know what that stuff is doing.

This is an `.Rmd` file. It is plain text with special features. Any time you write just like this, it will be compiled to normal text in the website. If you put a \# in front of your text, it will create a top level-header.

# My first post

Last compiled: `r Sys.Date()`

Notice that whatever you define as a top level header, automatically gets put into the table of contents bar on the left. 

## Second level header

You can add more headers by adding more hashtags. These won't be put into the table of contents

### third level header

Here's an even lower level header

# My second post (note the order)

Last compiled: `r Sys.Date()`

I'm writing this tutorial going from the top down. And, this is how it will be printed. So, notice the second post is second in the list. If you want your most recent post to be at the top, then make a new post starting at the top. If you want the oldest first, do, then keep adding to the bottom

# Adding R stuff

So far this is just a blog where you can write in plain text and serve your writing to a webpage. One of the main purposes of this lab journal is to record your progress learning R. The reason I am asking you to use this process is because you can both make a website, and a lab journal, and learn R all in R-studio. This makes everything really convenient and in the same place. 

So, let's say you are learning how to make a histogram in R. For example, maybe you want to sample 100 numbers from a normal distribution with mean = 0, and standard deviation = 1, and then you want to plot a histogram. You can do this right here by using an r code block, like this:

```{r}
# 1.0 Load libraries ----

library(tidyverse)
#  library(tibble)    --> is a modern re-imagining of the data frame
#  library(readr)     --> provides a fast and friendly way to read rectangular data like csv
#  library(dplyr)     --> provides a grammar of data manipulation
#  library(magrittr)  --> offers a set of operators which make your code more readable (pipe operator)
#  library(tidyr)     --> provides a set of functions that help you get to tidy data
#  library(stringr)   --> provides a cohesive set of functions designed to make working with strings as easy as possible
#  library(ggplot2)   --> graphics

## ── Attaching packages ──────────────────────────────── tidyverse 1.3.0 ──
## ✓ ggplot2 3.3.0     ✓ purrr   0.3.4
## ✓ tibble  3.0.1     ✓ dplyr   0.8.5
## ✓ tidyr   1.0.2     ✓ stringr 1.4.0
## ✓ readr   1.3.1     ✓ forcats 0.5.0
## ── Conflicts ───────────────────────────────────────── tidyverse_conflicts() ──
## x dplyr::filter() masks stats::filter()
## x dplyr::lag()    masks stats::lag()

# Excel Files
library(readxl)

# 2.0 Importing Files ----
# A good convention is to use the file name and suffix it with tbl for the data structure tibble
bikes_tbl      <- read_excel(path = "C:/data-science/00_data/01_bike_sales/01_raw_data/bikes.xlsx")
orderlines_tbl <- read_excel("C:/data-science/00_data/01_bike_sales/01_raw_data/orderlines.xlsx")

# Not necessary for this analysis, but for the sake of completeness
bikeshops_tbl  <- read_excel("C:/data-science/00_data/01_bike_sales/01_raw_data/bikeshops.xlsx")

# Method 1: Print it to the console
orderlines_tbl
# A tibble: 15,644 x 7
##    ...1  order.id order.line order.date          customer.id product.id quantity
##    <chr>    <dbl>      <dbl> <dttm>                    <dbl>      <dbl>    <dbl>
##  1 1            1          1 2015-01-07 00:00:00           2       2681        1
##  2 2            1          2 2015-01-07 00:00:00           2       2411        1
##  3 3            2          1 2015-01-10 00:00:00          10       2629        1
##  4 4            2          2 2015-01-10 00:00:00          10       2137        1
##  5 5            3          1 2015-01-10 00:00:00           6       2367        1
##  6 6            3          2 2015-01-10 00:00:00           6       1973        1
##  7 7            3          3 2015-01-10 00:00:00           6       2422        1
##  8 8            3          4 2015-01-10 00:00:00           6       2655        1
##  9 9            3          5 2015-01-10 00:00:00           6       2247        1
## 10 10           4          1 2015-01-11 00:00:00          22       2408        1
## # … with 15,634 more rows

# Method 2: Clicking on the file in the environment tab (or run View(orderlines_tbl)) There you can play around with the filter.

# Method 3: glimpse() function. Especially helpful for wide data (data with many columns)
glimpse(orderlines_tbl)
## Rows: 15,644


# Chaining commands with the pipe and assigning it to order_items_joined_tbl
bike_orderlines_joined_tbl <- orderlines_tbl %>%
  left_join(bikes_tbl, by = c("product.id" = "bike.id")) %>%
  left_join(bikeshops_tbl, by = c("customer.id" = "bikeshop.id"))

# Examine the results with glimpse()
bike_orderlines_joined_tbl %>% glimpse()
## Rows: 15,644
## Columns: 19


bike_orderlines_joined_tbl %>% 
  select(category) %>%
  filter(str_detect(category, "^Mountain")) %>% 
  unique()
## # A tibble: 10 x 1

# 5.0 Wrangling Data ----
# All actions are chained with the pipe already. You can perform each step separately and use glimpse() or View() to validate your code. Store the result in a variable at the end of the steps.
bike_orderlines_wrangled_tbl <- bike_orderlines_joined_tbl %>%
  # 5.1 Separate category name
  separate(col    = location,
           into   = c("city", "state"),
           sep    = ",") %>%
  
  # 5.2 Add the total price (price * quantity) 
  # Add a column to a tibble that uses a formula-style calculation of other columns
  mutate(total.price = price * quantity) %>%
  
  # 5.3 Optional: Reorganize. Using select to grab or remove unnecessary columns
  # 5.3.1 by exact column name
  select(-...1, -gender) %>%
  
  # 5.3.2 by a pattern
  # You can use the select_helpers to define patterns. 
  # Type ?ends_with and click on Select helpers in the documentation
  select(-ends_with(".id")) %>%
  
  # 5.3.3 Actually we need the column "order.id". Let's bind it back to the data
  bind_cols(bike_orderlines_joined_tbl %>% select(order.id)) %>% 
  
  # 5.3.4 You can reorder the data by selecting the columns in your desired order.
  # You can use select_helpers like contains() or everything()
  select(order.id, contains("order"), contains("model"), contains("location"),
         price, quantity, total.price,
         everything()) %>%
  
  # 5.4 Rename columns because we actually wanted underscores instead of the dots
  # (one at the time vs. multiple at once)
  rename(bikeshop = name) %>%
  set_names(names(.) %>% str_replace_all("\\.", "_"))

# 6.0 Business Insights ----
# 6.1 Sales by Year ----

library(lubridate)
# Step 1 - Manipulate
sales_by_year_tbl <- bike_orderlines_wrangled_tbl %>%
 
  # Select columns
  select(order_date, total_price) %>%

  # Add year column
  mutate(year = year(order_date)) %>%
 
  # Grouping by year and summarizing sales
  group_by(year) %>% 
  summarize(sales = sum(total_price)) %>%
 
  # Optional: Add a column that turns the numbers into a currency format 
  # (makes it in the plot optically more appealing)
  # mutate(sales_text = scales::dollar(sales)) <- Works for dollar values
  mutate(sales_text = scales::dollar(sales, big.mark = ".", 
                                     decimal.mark = ",", 
                                     prefix = "", 
                                     suffix = " €"))

sales_by_year_tbl
## # A tibble: 5 x 3

# 6.1 Sales by Year ----

  # Step 2 - Visualize
  sales_by_year_tbl %>%
  
  # Setup canvas with the columns year (x-axis) and sales (y-axis)
  ggplot(aes(x = year, y = sales)) +
  
  # Geometries
  geom_col(fill = "#2DC6D6") + # Use geom_col for a bar plot
  geom_label(aes(label = sales_text)) + # Adding labels to the bars
  geom_smooth(method = "lm", se = FALSE) + # Adding a trendline

  # Formatting
  # scale_y_continuous(labels = scales::dollar) + # Change the y-axis. 
  # Again, we have to adjust it for euro values
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  labs(
    title    = "Revenue by year",
    subtitle = "Upward Trend",
    x = "", # Override defaults for x and y
    y = "Revenue"
  )

# 6.2 Sales by Year and Category ----
# Step 1 - Manipulate
sales_by_year_cat_1_tbl <- bike_orderlines_wrangled_tbl %>%
 
  # Select columns and add a year
  select(order_date, total_price, state) %>%
  mutate(year = year(order_date)) %>%

  # Group by and summarize year and main catgegory
  group_by(year, state) %>%
  summarise(sales = sum(total_price)) %>%
  ungroup() %>%

  # Format $ Text
  mutate(sales_text = scales::dollar(sales, big.mark = ".", 
                                     decimal.mark = ",", 
                                     prefix = "", 
                                     suffix = " €"))

sales_by_year_cat_1_tbl  
## # A tibble: 25 x 4

# Step 2 - Visualize
sales_by_year_cat_1_tbl %>%

  # Set up x, y, fill
  ggplot(aes(x = year, y = sales, fill = state)) +

  # Geometries
  geom_col() + # Run up to here to get a stacked bar plot

  # Facet
  facet_wrap(~ state) +

  # Formatting
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  labs(
    title = "Revenue by year and state",
    subtitle = "Each product category has an upward trend",
    fill = "State" # Changes the legend name
  )
```

When you knit this R Markdown document, you will see that the histogram is printed to the page, along with the R code. This document can be set up to hide the R code in the webpage, just delete the comment (hashtag) from the cold folding option in the yaml header up top. For purposes of letting yourself see the code, and me see the code, best to keep it the way that it is. You'll learn that all of these things and more can be customized in each R code block.